---
title: "lipd2neotoma Demo"
author: "Dave Edge"
date: "2022-12-05"
output: html_document
---

```{r setup, include=FALSE}
library(lipdR)
knitr::opts_chunk$set(echo = TRUE)
```

## This is a demonstration of the new `lipd2neotoma()` function

### I will use a neotoma dataset, called Bambili 2, for this demonstration

First, we will load this record and convert into lipd format

```{r}
B <- neotoma2::get_sites(sitename = "Bambili 2")
D <- neotoma2::get_downloads(B)
L <- neotoma2lipd(D)
```

### LiPD

Let's see a summary of the record in LiPD

```{r warning=FALSE}
summary(L)
```


### Neotoma

Now we will convert back to neotoma and observe the changes

```{r}
N <- lipd2neotoma(L)

D@sites[[1]]
N

```

The site printout looks good

### Sample datum

```{r}
N <- lipd2neotoma(L)

D@sites[[1]]@collunits@collunits[[1]]@datasets@datasets[[1]]@samples@samples[[1]]@datum
N@collunits@collunits[[1]]@datasets@datasets[[1]]@samples@samples[[1]]@datum
```

note that `D` is the original neotoma dataset and `N` is the new one
`N` is a `site` rather than a `sites`

Aside from some loss in the `variablename`, these data frames look identical
Improved `variablename` fidelity is possible with alteration to the `neotoma2lipd()` function

What's the difference between element and elementtype?

Let's jump to the last `datum` to see if this fidelity holds

```{r}
length(D@sites[[1]]@collunits@collunits[[1]]@datasets@datasets[[1]]@samples@samples)
D@sites[[1]]@collunits@collunits[[1]]@datasets@datasets[[1]]@samples@samples[[140]]@datum
N$collunits@collunits[[1]]@datasets@datasets[[1]]@samples@samples[[140]]@datum
```

again, we see some minor changes to `variablename`, but otherwise the data frame is identical

### Chronologies

Let's have a look at the chronology data

```{r}
D@sites[[1]]@collunits@collunits[[1]]@chronologies@chronologies[[1]]@chroncontrols
N$collunits@collunits[[1]]@chronologies@chronologies[[1]]@chroncontrols
```

we can see two significant changes right off, different sorting and more columns

I could not identify the sorting of the original chronology table to follow

The new table contains the layer thickness data and age uncertainty
This could be changed to conform to the orignal style if desired

### Outstanding Issues

Collection unit print doesn't work 

```{r error=TRUE}
D@sites[[1]]@collunits@collunits[[1]]
N$collunits@collunits[[1]]
```

Some of the chronology metadata isn't retained

```{r error=TRUE}
D@sites[[1]]@collunits@collunits[[1]]@chronologies@chronologies[[1]]
N$collunits@collunits[[1]]@chronologies@chronologies[[1]]
```


