---
title: "LiPD and Neotoma interoperability"
author: "lipdR authors"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{LiPD and Neotoma interoperability}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output: rmarkdown::html_vignette
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = FALSE)
```

lipdR can convert between the [LiPD](https://lipdverse.org) and
[Neotoma](https://www.neotomadb.org) ecosystems in both directions. This makes it
possible to pull records out of the Neotoma Paleoecology Database, work with them
using the LiPD tools, and push results back.

The conversion functions require the suggested package `neotoma2` (available on CRAN)
and, for spatial handling, `sf`:

```{r install}
install.packages(c("neotoma2", "sf"))
```

The code chunks below are not evaluated when the vignette is built, because they make
live calls to the Neotoma API.

## Neotoma to LiPD

Download a site from Neotoma and convert it into a LiPD object with `neotoma2lipd()`:

```{r neo2lipd}
library(lipdR)

# Retrieve a site and its data from Neotoma
sites <- neotoma2::get_sites(sitename = "Bambili 2")
downloads <- neotoma2::get_downloads(sites)

# Convert to a LiPD object
L <- neotoma2lipd(downloads)

summary(L)
```

The resulting object is a standard LiPD object: it can be validated, written to a
`.lpd` file, flattened to a time series, and passed to the rest of the ecosystem
(geoChronR, actR, compositeR, and so on).

```{r use-lipd}
validLipd(L)
ts <- extractTs(L)
writeLipd(L, "~/Downloads/")
```

## LiPD to Neotoma

The reverse conversion, `lipd2neotoma()`, builds a Neotoma `site` object from a LiPD
object:

```{r lipd2neo}
L <- readLipd("path/to/a/file.lpd")

site <- lipd2neotoma(L)

# The result is a neotoma2 'site' object
neotoma2::samples(site)
```

## Round-tripping

Because both directions are supported, a LiPD object can be pushed to Neotoma and back:

```{r roundtrip}
site <- lipd2neotoma(L)
L2   <- neotoma2lipd(site)
```

Core metadata (coordinates, publications, measurement data) is preserved across the
round-trip, though the two data models are not identical, so some ecosystem-specific
fields may not survive.
