---
title: "LiPD Queries"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{createColumn}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, width.out='100%'}

options(width = 500)
knitr::opts_chunk$set(
  error=TRUE,
  collapse = TRUE,
  comment = "#>"
)
```

The `queryGet()` function in `lipdR` allows users to download LiPD files based on various filter parameters

The filtering relies on a query table, which holds metadata for all time series available on LiPDverse

This query table is built in to, and loaded with lipdR

```{r echo=T, results='hide'}
devtools::load_all()
```

Let's get a sense for what the table holds

```{r}
dim(queryTable)
names(queryTable)
```

That's a lot of rows! There's a row for every time series in the LiPDverse database

If you have used the LiPD format, you will recognize some of these names

Let's try a simple query

We'll set grab.data to FALSE since we aren't ready to download anything yet

```{r}
queryGet(variable.name = c("δ18O"),
               grab.data = FALSE)
```
huh, there must be more Oxygen-18 time series

Let's have a look at the unique paleoData `variable.name`

We'll just look at the first 250, there are quite a few

```{r}
unique(queryTable$paleoData_variableName)[1:250]
```

There's quite a few notation styles, but they all have "18o" in common, so let's try that

```{r}
queryGet(variable.name = c("18o"),
               grab.data = FALSE)
```

Now that is a lot more data

Some query parameters will require considering what all the possible results have in common

In other cases, we can simply input a vector with several possible options

For the `variable.name` parameter, multiple entries are combined with "OR" logic, 
so more entries will generally pull more datasets

```{r}
queryGet(variable.name = c("δ18O", "d18O"),
               grab.data = FALSE)
```

This gets us almost the same result, but we see the simplified filter is still pulling one more dataset

Also note that capitalization makes no difference here

The `archive.type` filter works similarly,
let's look at our options

```{r}
unique(queryTable$archiveType)
```
Note that we have distinct options: "Lake", "LakeSediment", "LakeDeposits", "LakeDeposit", and "Lake Sediment"

We can grab the `archive.type` names and search for all of them like this

```{r}
queryGet(archive.type = unique(queryTable$archiveType)[c(2,3,5,6,39)],
               grab.data = FALSE)
```
Or we can use the simplification strategy again

```{r}
queryGet(archive.type = c("lake"),
               grab.data = FALSE)
```

As we saw last time, the results are similar

We can also filter based on publications: Author, DOI, Title, etc. using `pub.info`

```{r}
queryGet(pub.info = c("10.1029/2008PA0016718K7NFLXQ1"),
               grab.data = FALSE)
```

This query can be a little slow if you don't narrow the results with another parameter first

Let's narrow our region of interest

There are four different parameters used for this: `coord`, `country`, `continent`, and `ocean`

Let's grab all the North American datasets

```{r}
queryGet(continent = "North America",
               grab.data = FALSE)
```

Let's grab just those from Mexico

```{r}
queryGet(country = c("Mexico"),
         grab.data = FALSE)
```
Note that the country and contient filters are not filtered based on LiPD content. A function in R uses the coordinates associated with the datasets to associate them with countries and the results can be unreliable near country borders

Again, we can see all of the options for `country` and `continent` with `unique()`

We can also use latitude and longitude directly with a bounding box

```{r}
queryGet(coord = c(0,90,-180,-110),
         grab.data = FALSE)
```
We can limit this to only marine data by setting `ocean` to `TRUE`

```{r}
queryGet(coord = c(0,90,-180,-110),
         ocean = TRUE,
         grab.data = FALSE)
```
The `ocean` parameter works on the same basis as the `country` and `continent` parameters and can be unreliable in coastal areas

We can also grab all the data from a `compilation`, such as the Western North America (WNAm)

```{r}
queryGet(compilation = c("wnam"), grab.data = FALSE)
```

or multiple compilations

```{r}
queryGet(compilation = c("wnam", "temp12k"), grab.data = FALSE)
```
the `compilation` filter uses "OR" logic

Let's try pulling datasets based on their `seasonality`

We can pull summer, commonly defined as June, July, and August, for the Northern Hemisphere

`seasonality` input is taken as a list

```{r}
queryGet(continent = "North America",
               seasonality = list("June", "July", "August"),
               grab.data = FALSE)
```

Okay, so this must not be a good choice of format, let's see how LiPD authors define their seasons

```{r}
unique(queryTable$interpretation1_seasonality)

```

It looks like numeric months are most common. 
Season names, "warm"/"cold", and series of first-letter abbreviations (ie. JJA) are all common.

Knowing this, let's try again

Items within a single list are treated as linked by "AND", so an input of list("June", "July", "August")
would filter for season data with ALL of these months

Multiple lists are treated as linked by "OR", such that list(list("June"), list("July"), list("August))
would filter for season data with ANY of these months

Let's try to get all of the summer datasets by entering a few different notations

```{r}
queryGet(continent = "North America",
               seasonality = list(list("6", "7", "8"), list("summer"), list("JJA")),
               grab.data = FALSE)
```

This returns quite a few datasets

From our look at all the unique `seasonality` entries, we can see that this probably includes a lot of annual data too

Let's exclude the annual and winter datasets by using `season.not`

The input for for `season.not` works the same as `seasonality`
```{r}
queryGet(continent = "North America",
               seasonality = list(list("6", "7", "8"), list("summer"), list("JJA")),
               season.not = list(list("annual"), list("December"), list("12","1","2"), list("winter"), list("cold")),
               grab.data = FALSE)
```

Now we've narrowed it down to summer-specific datasets

Let's look at interpretation variables now. These are the climate variables that may serve as a target for the proxy time series available

These variables are expressed in two different formats: interpretation variable and interpretation detail

Each of these variables has four possible interpretation slots

```{r}
unique(queryTable$interp_Vars)

unique(queryTable$interp_Details)
```

Let's look at some marine `interp.vars` in the northeast Pacific

```{r}
queryGet(coord = c(0,90,-180,-110),
               ocean = TRUE,
               interp.vars =  c("SST", "upwelling", "SSS"),
               grab.data = FALSE)
```

That gives us just a few datasets

Perhaps we'll have more luck with `interp.details`, which is more standardized

```{r,warning=FALSE}
queryGet(coord = c(0,90,-180,-110),
               ocean = TRUE,
               interp.details = c("sea@surface", "elNino"),
               grab.data = FALSE)

```

let's see if we grab more using both

These inputs combine with "OR" logic, so we may gather more datasets by using both parameters

```{r}
queryGet(coord = c(0,90,-180,-110),
               ocean = TRUE,
               interp.details = c("sea@surface", "elNino"),
               interp.vars =  c("SST", "upwelling", "SSS"),
               grab.data = FALSE)
```

looks like we get a few extra datasets with this approach

Now that we know how to use our filters, let's go for a strict filter

We'll look for marine archives in the northeast Pacific, with interpretations
related to marine climate variables in the summer months only

```{r}
queryGet(coord = c(0,90,-180,-110),
               archive.type = c("marine", "ocean"),
               ocean = TRUE,
               interp.details = c("sea@surface", "elNino"),
               interp.vars =  c("SST", "upwelling", "SSS"),
               seasonality = list(list("6", "7", "8"), list("summer"), list("JJA")),
               season.not = list(list("annual"), list("December"), list("12","1","2"), list("winter"), list("cold")),
               grab.data = FALSE)
```

To fine-tune your query, set verbose to TRUE to see which parameters have what effect on filtering

We'll narrow the results further by adding author names to find within the publication info

```{r}
queryGet(coord = c(0,90,-180,-110),
               archive.type = c("marine", "ocean"),
               ocean = TRUE,
               interp.details = c("sea@surface", "elNino"),
               interp.vars =  c("SST", "upwelling", "SSS"),
               seasonality = list(list("6", "7", "8"), list("summer"), list("JJA")),
               season.not = list(list("annual"), list("December"), list("12","1","2"), list("winter"), list("cold")),
               grab.data = FALSE,
               pub.info = c("mix", "caissie"),
               verbose = TRUE)
```

When you're satisfied with the query results, change grab.data to TRUE and assign the function to download the LiPD datasets

```{r}
D <- queryGet(coord = c(0,90,-180,-110),
               archive.type = c("marine", "ocean"),
               ocean = TRUE,
               interp.details = c("sea@surface", "elNino"),
               interp.vars =  c("SST", "upwelling", "SSS"),
               seasonality = list(list("6", "7", "8"), list("summer"), list("JJA")),
               season.not = list(list("annual"), list("December"), list("12","1","2"), list("winter"), list("cold")),
               grab.data = TRUE,
               pub.info = c("mix", "caissie"),
               verbose = FALSE,
              )
```