---
title: "lipdKeystandardization with ISO2k"
author: "Dave Edge"
date: "2023-01-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
devtools::load_all()
```

###Load Data

First we grab the ISO2k data and the standardization tables

```{r}
load("C:/Users/dce25/Downloads/iso2k1_0_1.RData")
standardTables <- readRDS("C:/Users/dce25/Downloads/standardTables.RDS")
```

###Check for valid `archiveType`

`isValidKey()` checks to see if all terms for a given key are valid lipd terms

```{r}
# print(TS)
isValidKey(TS, "archiveType")
```

the function returns a data frame showing invalid terms


Now we can standardize the invalid terms by looking for synonyms

`standardizeKey()` looks for matches from previous lipd entries and PaST terms which are know to be synonyms of valid lipd terms

```{r, warning=FALSE}
TSnew <- standardizeKey(TS, "archiveType")
```

the function was able to find matches for some, but not all invalid terms

The function returns the TS object (now `TSnew`) with corrections where possible

The user is given the option to make the suggested corrections or abort


###Check for valid `variableName`

Same process with a different key

```{r}
# print(TS)
isValidKey(TS, "paleoData_variableName")
```

Wow, there a lot of invalid terms here

```{r}
TSnew <- standardizeKey(TS, "paleoData_variableName")
```

And this time we do have synonyms allowing for automated corrections to all invalid terms




