---
title: "lipd Values standardization with ISO2k"
author: "Dave Edge"
date: "2023-01-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
devtools::load_all()
```

###Load Data

First we grab the ISO2k data and the standardization tables

```{r}
load("C:/Users/dce25/Downloads/iso2k1_0_1.RData")
```

First, update the metadata

```{r}
TS1 <- updateMetaDataFromStandardTables(TS, "paleoData_units")
```

This function produces an updated TS object as well as a data frame showing the changes

`standardizeValue()` looks for synonyms and replaces them with valid lipd names

```{r}
TS2 <- standardizeValue(lipdTS=TS1$TS, key="paleoData_units")
```

#This function also produces a data frame showing any changes

A quick sanity check

```{r}
isValidValue(TS2$TS, "paleoData_units")
TS[[2]]$paleoData_units
TS2$TS[[2]]$paleoData_units
```

#Looks good

Finally we can update either `paleoData_notes` or `notes` with the changes made

```{r}
TS1 <- lipdR:::updateNotes(lipdTS=TS2$TS,
                           key = "paleoData_units",
                           metadataChangesDF=TS1$ChangesDF,
                           standardizeSynonymDF=TS2$synonymDF$synonymDF)
```

Let's have a look at some examples

unchanged
```{r}
TS[[1]]$paleoData_notes
TS1[[1]]$paleoData_notes
```

#Changed a peice of metadata

```{r}
TS[[2]]$paleoData_notes
TS1[[2]]$paleoData_notes
```


#Changed metadata and standardized value

```{r}
TS[[240]]$paleoData_notes
TS1[[240]]$paleoData_notes
```

Now let's try this for a more complicated key


```{r}
TS1 <- updateMetaDataFromStandardTables(TS, "interpretation_seasonality")
TS1
```

There is no metadata to update this time

Let's try standardizing

```{r}
TS2 <- standardizeValue(TS, "interpretation_seasonality")
```

We see a long string of output here, reflecting the 7 unique interpretation_seasonality keys

Let's have a look now

```{r}
a2 <- isValidValue(lipdTS=TS2$TS, key="interpretation_seasonality")
```

Cool, that took care of all of them

An example of values changed

```{r}
TS[[1851]]$interpretation5_seasonality
TS2$TS[[1851]]$interpretation5_seasonality
```


#Finally, the notes
#This is set up to be run once per interpretation, although the automation could be added
```{r}
TS1 <- lipdR:::updateNotes(lipdTS=TS2$TS,
                           standardizeSynonymDF=TS2$synonymDF[[1]]$synonymDF)
```

#before and after

```{r}
TS[[328]]$notes
TS1[[328]]$notes
```

#for interp 5

```{r}
TS1 <- lipdR:::updateNotes(lipdTS=TS2$TS,
                           standardizeSynonymDF=TS2$synonymDF[[5]]$synonymDF)
```

#before and after

```{r}
TS[[1851]]$notes
TS1[[1851]]$notes
```
