---
title: "LiPD Summaries"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{createColumn}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include = FALSE, width.out='100%'}
devtools::load_all("C:\\Users\\dce72\\Documents\\LiPDsandbox\\lipdR-Summarize")
#library(lipdR)
#source("C:\\Users\\dce72\\Documents\\LiPDsandbox\\lipdR-Summarize\\R\\summary.R")
#source("C:\\Users\\dce72\\Documents\\LiPDsandbox\\lipdR-Summarize\\R\\is_blank.R")
#source("C:\\Users\\dce72\\Documents\\LiPDsandbox\\lipdR-Summarize\\R\\cbind_NA.R")
#source("C:\\Users\\dce72\\Documents\\LiPDsandbox\\lipdR-Summarize\\R\\s3.R")
options(width = 500)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r xaringanExtra-clipboard, echo=FALSE}
xaringanExtra::use_clipboard()
```

## Data
### Load LiPD files from lipdverse.org
##### From the Pages 2k collection, let's look at the Moose Lake dataset
##### You can download the file first or simply use the url in readLipd()

```{r echo=T, results='hide'}
L <- readLipd("https://lipdverse.org/data/MD6jkgwSxsq0oilgYUjM/1_0_0//Arc-MooseLake.Clegg.2010-ensemble.lpd")

```
## Single LiPD
### See a detailed summary of a single LiPD
##### The single lipd summary provides detail on the location, publications(s), paleo data, chron data, and model ensembles

```{r}

summary(L)

```
## Multi LiPD

### Let's use a large library of lipd files

```{r echo = T, results = 'hide'}
D <- readLipd("http://lipdverse.org/testData/testData.zip")

```


### Get a summary of all LiPDs in a multi-lipd object

##### We can choose the prefered age units. Let's return time values in years AD if available

##### Leaving return.table set to 'FALSE' will print the table to the screen

##### Let's print the details of the first 20 datasets

```{r}

multiLipdSummary(multi.lipd=D, print.length=20, time.units="AD", return.table=FALSE)

```
### Save the summary table

##### Sometimes it is useful to save the table, so it can be sorted and further investigated

##### Simply change return.table to 'TRUE' and save the summary

```{r}

MLSummary <- multiLipdSummary(D, time.units="AD", return.table = TRUE)

```
### The saved table can now be interrogated

##### For example, we can sort the table to find datasets with age ensembles

```{r}
MLSummary[order(MLSummary$NumEns, decreasing=TRUE),]
```


## LiPD TS Objects

### Format data as a tibble LiPD TS

##### extractTs() creates a time series from a multi-lipd object
##### This allows for easier querying and data analysis

```{r}

TS1 <- extractTs(D)
tibTS <- ts2tibble(TS1)


```

### View the dimensions of the tibble TS object

```{r}
dim(tibTS)
```

### Print summary of LiPD TS

##### The lipdTSSummary() function accepts both lipd-ts (TS1) and lipd-tib-ts (tibTS) objects

##### The summary returns the dimensions of the TS object, details of the time units, a time interval covered by all varibales (if possible), and a summary table

##### Let's limit the table to show details of the first 10 rows only

```{r}
lipdTSSummary(ts.object=tibTS, print.length = 10, return.table = FALSE)

```
### Add columns to summary

##### The summary table can be expanded to include any variables of the TS object
##### The names() function shows the options (here we see the first 5)
##### Let's add several
##### The summary function handles numeric and character variables differently

```{r}
names(tibTS)[1:5]

lipdTSSummary(tibTS, print.length = 10, return.table = FALSE, add.variable = c("archiveType", "paleoData_meanValue12k","geo_continent", "geo_elevation", "agesPerKyr", "dataSetName"))

```



### Save summary table and view a specific subset

##### Saving the summary table allows for further investigation

##### Let's subset the continental United States (roughly)

##### Let's choose only those datasets covering the early Holocene

##### Now, let's pull all the climate interpretation info

##### We can now have a closer look to see which of these data are appropriate

```{r}
#Continental U.S. bounds: -125	-70 30	50
tibTSus <-tibTS[tibTS$geo_latitude > 30 & tibTS$geo_latitude < 50,]
tibTSus <-tibTSus[tibTSus$geo_longitude > -125 & tibTSus$geo_longitude < -70,]
#Early Holocene
tibTSus <-tibTSus[apply(tibTSus, 1, function(x) max(x$age)>11700),]
tibTSus <-tibTSus[apply(tibTSus, 1, function(x) min(x$age)>100),]
#grab interpreation info
intepIndices <- grep("interpretation", names(tibTSus))
hasInterp <- intepIndices[apply(tibTSus[,intepIndices], 2, function(x) sum(!is.na(x))>0)]

#save summary for deeper exploration
USsummary <- lipdTSSummary(tibTSus, print.length = 10, return.table = TRUE, add.variable = names(tibTSus[,hasInterp]))
USsummary
```

