---
title: "Exploring LiPD data using lipdR summary functions"
author: "David Edge"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Summarize LiPD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



```{r setup, include = FALSE, fig.width=20}
devtools::load_all("C:\\Users\\dce72\\Documents\\LiPDsandbox\\lipdR-Summarize")
options(width = 150)
library(xaringanExtra)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<style type="text/css">
div.main-container {
  max-width: 1600 ox !important;
  margin: auto;
}
</style>

```{r xaringanExtra-clipboard, echo=FALSE}
xaringanExtra::use_clipboard()
```

## Data
### Load LiPD files from lipdverse.org
```{r echo=T, results='hide'}

D <- readLipd("http://lipdverse.org/testData/testData.zip")

```
## Single LiPD
### See a detailed summary of a single LiPD
```{r}

L1 <- D[[1]]
lipdSummary(L1)

```
## Multi LiPD
### Get a summary of all LiPDs in the directory, details of first 20
### Return time values in "AD" units if available
```{r}

lipdDirSummary(D, printLen=20, timeUnits="AD", retTable = FALSE)

```
### save the summary table

```{r}

dirSummary <- lipdDirSummary(D, printLen=20, timeUnits="AD", retTable = TRUE)

```
### sort table to find age ensembles

```{r}
dirSummary[order(dirSummary$NumEns, decreasing=TRUE),]
```


## LiPD TS Objects
### Format data as a tibble LiPD TS
```{r}

TS1 <- extractTs(D)
tibTS <- ts2tibble(TS1)

```

### View dimensions of tibble object
```{r}
dim(tibTS)
```

### Print summary of LiPD TS, detail first 10 only
```{r}

lipdTSSummary(tibTS, printLen = 10, retTable = FALSE)

```
### Add columns to summary
```{r}
lipdTSSummary(tibTS, printLen = 10, retTable = FALSE, addCol = c("archiveType", "paleoData_meanValue12k"))
```



### Save summary table and view a specific subset
```{r}

TSsummary <- lipdTSSummary(tibTS, printLen = 10, retTable = TRUE)

tibTS1 <- TSsummary[TSsummary$dsName == "165_1002C.Herbert.2000",]
tibTS1

```
### Subset the TS object and view a summary

```{r}

tibTS2 <- tibTS[grep("tree", tibTS$archiveType),]
lipdTSSummary(tibTS2, printLen = 50)

```




